---
title: An R Markdown document converted from "36_PopDat.ipynb"
output: html_document
---

# Covariates for PRS analyses

---

## Libraries

```{r}
library(readr)
library(dplyr, warn.conflicts = FALSE)
library(tidyr, warn.conflicts = FALSE)
library(purrr)
```

---

## Baseline data

```{r}
tab_name <- "~/projects/DVA/Data/UKBiobank/ukb_57232/data_files/41806.tab"
```

Locating columns needed:

```{r}
tab_vars <- tribble(
    ~colid, ~colnam, ~coltyp,
    ## Basic
    "f.eid", "eid", "n",
    "f.53.0.0", "date0", "D",
    "f.31.0.0", "sex", "n",
    "f.21003.0.0", "age", "n",
    "f.4080.0.0", "sbp_01", "n",
    "f.4080.0.1", "sbp_02", "n",
    "f.30700.0.0", "creatinine", "n",
    "f.30760.0.0", "hdl", "n",
    "f.30780.0.0", "ldl", "n",
    "f.30870.0.0", "tg", "n",
    "f.30750.0.0", "hba1c", "n",
    "f.1239.0.0", "current_smoker", "n",
    "f.22006.0.0", "caucasian", "n",
    "f.22001.0.0", "genetic_sex", "n",
    "f.22019.0.0", "sexchr_aneuploidy", "n",
    "f.22020.0.0", "usedingpc", "n",
    "f.2724.0.0", "menopause", "n"
) %>%
    bind_rows(
        tibble(
            colid = paste0("f.22009.0.", 1:10),
            colnam = paste0("gpc", 1:10),
            coltyp = "n"
        )
    )
```

```{r}
tab_vars <- paste("head -1", tab_name, "| tr '\t' '\n'") %>%
    pipe %>%
    read_table(col_names = "colid", col_types = "c") %>%
    ## Position of columns in table
    mutate(colpos = row_number()) %>%
    ## Extracting position and name of variables of interest
    inner_join(tab_vars, by = "colid")
head(tab_vars)
```

```{r}
column_cmd <- paste("cut -f", paste(tab_vars$colpos, collapse = ","), tab_name)
cat(column_cmd)
```

```{r}
crude_dat <- read_tsv(
    file = pipe(column_cmd),
    ## Replacing column header with variable names
    skip = 1, col_names = tab_vars$colnam,
    col_types = paste(tab_vars$coltyp, collapse = "")
)
head(crude_dat)
```

```{r}
recoded_dat <- crude_dat %>%
    filter(
        caucasian == 1,
        sex == genetic_sex,
        usedingpc == 1,
        is.na(sexchr_aneuploidy)
    ) %>%
    select(-c(caucasian, genetic_sex, usedingpc, sexchr_aneuploidy)) %>%
    transmute(
        eid, age, 
        sex = ifelse(sex == 0, "Female", "Male"),
        sbp = rowMeans(cbind(sbp_01, sbp_02), na.rm = TRUE),
        scr = creatinine / 88.42,
        alpha = case_when(sex == "Female" ~ -0.241, sex == "Male" ~ -0.302),
        kappa = case_when(sex == "Female" ~ 0.7, sex == "Male" ~ 0.9),
        creat_kappa = scr / kappa,
        minkappa = pmin(creat_kappa, 1), 
        maxkappa = pmax(creat_kappa, 1),
        minkappa_alpha = minkappa^alpha,
        maxkappa_exp = maxkappa^(-1.200),
        age_term = 0.9938^age,
        sex_term = case_when(sex == "Female" ~ 1.012, sex == "Male" ~ 1), 
        egfr = 142 * minkappa_alpha * maxkappa_exp * age_term * sex_term,
        hdl, hba1c,
        tchol = hdl + ldl + (tg/2.2),
        smoking = case_when(current_smoker %in% c(1,2) ~ 1,
                            current_smoker == 0 ~ 0,
                            TRUE ~ NaN),
        menopause = replace_na(1*(menopause == 1), 0),
        across(starts_with("gpc"))
    ) %>%
    select(-c(scr, alpha, kappa, creat_kappa, minkappa, maxkappa,
              minkappa_alpha, maxkappa_exp, age_term, sex_term))
head(recoded_dat)
nrow(recoded_dat)
```

```{r}
write_tsv(recoded_dat, "../data/recoded_dat.tsv")
```

## Non-cancer illness codes

```{r}
tab_vars <- tibble(
    colid = "f.eid",
    colnam = "eid"
) %>%
    bind_rows(
        data.frame(colid = paste0("f.20002.0.", 0:33),
                   colnam = paste("noncancercode", 1:34, sep = "_")),
        data.frame(colid = paste0("f.20009.0.", 0:33),
                   colnam = paste("noncancerage", 1:34, sep = "_"))
    )
tab_vars <- paste("head -1", tab_name, "| tr '\t' '\n'") %>%
    pipe %>%
    read_table(col_names = "colid", col_types = "c") %>%
    ## Position of columns in table
    mutate(colpos = row_number()) %>%
    ## Extracting position and name of variables of interest
    inner_join(tab_vars, by = "colid")
head(tab_vars)
```

```{r}
column_cmd <- paste("cut -f", paste(tab_vars$colpos, collapse = ","), tab_name)
cat(column_cmd)
```

```{r}
dx_dat <- read_tsv(
    file = pipe(column_cmd),
    ## Replacing column header with variable names
    skip = 1, col_names = tab_vars$colnam,
    ## All columns are numeric
    col_types = cols(.default = "n")
)
head(dx_dat)
```

Long format

```{r}
dx_dat2 <- dx_dat %>%
    pivot_longer(
        -eid,
        names_sep = "_",
        names_to = c(".value", NA)
    )
head(dx_dat2)
```

Joining with coding:

```{r}
dx_coding <- read_tsv("~/projects/DVA/Data/UKBiobank/ukb_57232/codings/coding6.tsv",
                      show_col_types = FALSE)
head(dx_coding)
```

```{r}
dx_dat2 <- dx_dat2 %>%
    left_join(dx_coding[, c("coding", "meaning")],
               by = c("noncancercode" = "coding"))
head(dx_dat2)
```

### Selection of diseases for adjustment

```{r}
selected_dx <- dx_dat2 %>%
    group_by(eid) %>%
    summarise(
        HT = any(meaning %in% c("hypertension", 
                                "essential hypertension"), 
                 na.rm = TRUE),
        CKD = any(meaning %in% c("renal/kidney failure", "renal failure requiring dialysis",
                                 "renal failure not requiring dialysis"),
                  na.rm = TRUE),
        T2D = (any(meaning %in% c("diabetes", "type 2 diabetes", "diabetic eye disease", 
                                  "diabetic neuropathy/ulcers"), na.rm = TRUE) &
               !any(meaning %in% c("type 1 diabetes"), na.rm = TRUE)),
        T2Dage = replace_na(
            sort(
                noncancerage[
                    meaning %in% c(
                        "diabetes", "type 2 diabetes", 
                        "diabetic eye disease", "diabetic neuropathy/ulcers"
                    ) & noncancerage > 0
                ]
            )[1],
            0
        ),
        T1D = any(meaning == "type 1 diabetes", na.rm = TRUE)
    ) %>%
    mutate(
        across(where(is.logical), ~.x * 1),
        T1D = ifelse(T2D == 1 & T2Dage < 35, 1, T1D),
        T2Dage = ifelse(T2D == 1 & T1D == 0, T2Dage, 0),
        T1D = NULL
    )
head(selected_dx)
```

## Medication data

```{r}
tab_vars <- tibble(
    colid = "f.eid",
    colnam = "eid",
) %>%
    bind_rows(
        data.frame(colid = paste0("f.20003.0.", 0:47),
                   colnam = paste("medcode", 1:48, sep = "_"))
    )
tab_vars <- paste("head -1", tab_name, "| tr '\t' '\n'") %>%
    pipe %>%
    read_table(col_names = "colid", col_types = "c") %>%
    ## Position of columns in table
    mutate(colpos = row_number()) %>%
    ## Extracting position and name of variables of interest
    inner_join(tab_vars, by = "colid")
head(tab_vars)
```

```{r}
column_cmd <- paste("cut -f", paste(tab_vars$colpos, collapse = ","), tab_name)
cat(column_cmd)
```

```{r}
meds_dat <- read_tsv(
    file = pipe(column_cmd),
    ## Replacing column header with variable names
    skip = 1, col_names = tab_vars$colnam,
    ## All columns are numeric
    col_types = cols(.default = "n")
)
head(meds_dat)
```

Long format

```{r}
meds_dat2 <- meds_dat %>%
    pivot_longer(
        -eid,
        names_sep = "_",
        names_to = c(".value", NA)
    )
head(meds_dat2)
```

Joining with coding:

```{r}
meds_coding <- read_tsv("~/projects/DVA/Data/UKBiobank/ukb_57232/codings/coding4.tsv",
                      show_col_types = FALSE)
head(meds_coding)
```

```{r}
meds_dat2 <- meds_dat2 %>%
    left_join(meds_coding,
               by = c("medcode" = "coding"))
head(meds_dat2)
```

Map of UK Biobank medication to ATC codes (available for some medications only):

```{r}
ATC_matches <- read_tsv("~/projects/DVA/Data/UKBiobank/ukb_57232/codings/ATC_matches.tsv", 
                        col_types = "ncc")
head(ATC_matches)
```

Selected codes to use:

- A10A - Insulin
- A10B - Diabetes drugs
- C01 - Cardiac therapy
- C02 - Antihypertensives
- C03 - Diuretics
- C07 - Beta blocking agents
- C08 - Calcium channel blockers
- C09 - RA system drugs
- C10 - Lipid lowering

```{r}
ATC_matches_sel <- ATC_matches %>%
    filter(grepl("^A10|^C01|^C02|^C03|^C07|^C08|^C09|^C10", ATC_codes)) %>%
    unique
head(ATC_matches_sel)
```

```{r}
selmeds <- meds_dat2 %>%
    left_join(ATC_matches_sel, by = c("medcode" = "ukb_code", 
                                       "meaning" = "med_name"))
head(selmeds)
```

Ocurrence of selected medication:

```{r}
selected_meds <- selmeds %>%
    group_by(eid) %>%
    summarise(
        Insulin = any(ATC_codes %in% c("A10A"), na.rm = TRUE),
        AntiDM = any(ATC_codes %in% c("A10B"), na.rm = TRUE),
        AntiHT = any(grepl("^C01|^C02|^C03|^C07|^C08|^C09", ATC_codes), na.rm = TRUE),
        LipidLower = any(grepl("^C10", ATC_codes), na.rm = TRUE)
    ) %>%
    mutate(across(-eid, ~.x * 1)) %>%
    right_join(meds_dat[,"eid"], by = "eid") %>%
    mutate(across(-eid, ~replace_na(.x, 0)))
head(selected_meds)
nrow(selected_meds)
```

## Diagnosis and medication data

```{r}
dxmed <- inner_join(selected_dx, selected_meds, by = "eid")
head(dxmed)
```

```{r}
write_tsv(dxmed, "../data/dxmed.tsv")
```

## MACE data from hospital and death records

### Ocurrences in hospital records

Hospital records - main table:

```{r}
hesin <- read_tsv(
    "~/projects/DVA/Data/UKBiobank/ukb_57232/hes_data/hesin.txt",
    guess_max = 50000, show_col_types = FALSE
)
print(data.frame(head(hesin)))
```

Hospital records - Diagnosis table:

```{r}
hesin_diag <- read_tsv(
    "~/projects/DVA/Data/UKBiobank/ukb_57232/hes_data/hesin_diag.txt", 
    skip = 1,
    col_types = "nnn-c-c-",
    col_names = c("eid", "ins_index", "arr_index", "diag_icd9", "diag_icd10")
)
head(hesin_diag)
```

Joining tables:

```{r}
hesin_join <- hesin %>%
    transmute(eid, ins_index, 
              event_date = coalesce(epistart, epiend, elecdate, admidate, epiend, disdate),
              event_date = lubridate::dmy(event_date)) %>%
    inner_join(hesin_diag) %>%
    select(-c(ins_index, arr_index)) %>%
    rename(ICD_9 = diag_icd9, ICD_10 = diag_icd10) %>%
    pivot_longer(c(ICD_9, ICD_10), names_to = "CODESYS", values_to = "CODE", values_drop_na = TRUE)
head(hesin_join)
```

### Occurrences from mortality data

Mortality data - date:

```{r}
mordate <- read_tsv("~/projects/DVA/Data/UKBiobank/ukb_57232/mortality/death.txt", show_col_types = FALSE)
head(mordate)
```

Mortality data - cause:

```{r}
morcause <- read_tsv("~/projects/DVA/Data/UKBiobank/ukb_57232/mortality/death_cause.txt", show_col_types = FALSE)
head(morcause)
```

Mortality - Join tables:

```{r}
mor <- inner_join(mordate, morcause) %>%
    transmute(eid, 
              event_date = lubridate::dmy(date_of_death),
              CODESYS = "ICD_10", CODE = cause_icd10)
head(mor)
```

### Joining occurrences

```{r}
hesmort_outcomes <- bind_rows(
    hesin_join, mor
)
head(hesmort_outcomes)
```

### Finding MACE events

MACE codes:

```{r}
mace_codes <- read_tsv("../data/MACE_CODES.tsv", show_col_types = FALSE)
head(mace_codes)
```

```{r}
outcome_lookup <- mace_codes %>%
    group_by(OUTCOME, CODESYS) %>%
    nest %>%
    mutate(
        data = map_chr(data, ~paste0("^", .x$CODE, collapse = "|"))
    ) %>%
    ungroup
outcome_lookup
```

```{r}
outcome_lookup_res <- outcome_lookup %>%
    transmute(
        outcome = OUTCOME,
        res = map2(
            CODESYS, data,
            ~hesmort_outcomes %>% 
                filter(CODESYS == .x, 
                       grepl(.y, CODE)) %>%
                select(eid, event_date)
        )
    ) %>%
    unnest(res) %>%
    unique
head(outcome_lookup_res)
```

### Joining all occurrences of MACE

```{r}
allocurrmace <- outcome_lookup_res %>%
    transmute(eid, outcome_value = 1, outcome_date = event_date)
head(allocurrmace)
```

### First and last date for all individuals

```{r}
censordate <- lubridate::ymd("2018-02-28")
```

```{r}
datedat <- mor %>%
    select(eid, lastdate = event_date) %>%
    group_by(eid) %>%
    slice_min(lastdate, with_ties = FALSE) %>%
    ungroup %>%
    right_join(crude_dat[,c("eid", "date0")]) %>%
    transmute(
        eid, date0, 
        lastdate = replace_na(lastdate, censordate),
        lastdate = if_else(lastdate > censordate, censordate, lastdate),
        lastdate = if_else(lastdate > date0 + lubridate::years(10), 
                           date0 + lubridate::years(10), lastdate)
    ) %>%
    drop_na
head(datedat)
nrow(datedat)
```

### Survival data including prior MACE

```{r}
survmacedat <- datedat %>%
    left_join(allocurrmace, by = "eid") %>%
    mutate(
        outcome_value = replace_na(outcome_value, 0),
        outcome_date = coalesce(outcome_date, lastdate),
        outcome_value = if_else(outcome_date > lastdate, 0, outcome_value),
        outcome_date = if_else(outcome_date > lastdate, lastdate, outcome_date),
        priorMACE_timeyrs = pmin(0, as.numeric(lubridate::interval(date0, outcome_date), "years")),
        outcome_value = ifelse(priorMACE_timeyrs < 0, 0, outcome_value),
        outcome_date = if_else(priorMACE_timeyrs < 0, lastdate, outcome_date),
        outcome_timeyrs = as.numeric(lubridate::interval(date0, outcome_date), "years")
    ) %>%
    unique %>%
    group_by(eid) %>%
    mutate(priorMACE_timeyrs = min(priorMACE_timeyrs)) %>%
    slice_min(outcome_timeyrs, with_ties = FALSE) %>%
    ungroup %>%
    mutate(priorMACE = 1 * (priorMACE_timeyrs < 0)) %>%
    select(eid, outcome_value, outcome_timeyrs, priorMACE, priorMACE_timeyrs)
head(survmacedat)
```

```{r}
write_tsv(survmacedat, "../data/survmacedat.tsv")
```

